Deep Multimodal Generative Models for Weakly-Supervised Learning

### 2017-11-18 paper/draft.tex
Added images + placeholder tables for MNIST... need to do for MultiMNIST and then call it a day

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage[final]{nips_2017}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{natbib}

@@

\begin{equation}
    J(x, y, \theta, \phi) = \mu (E_{q_{\phi_{x}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{x}}(z|x)), p_{\theta}(z)) + (1 - \mu) (E_{q_{\phi_{y}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{y}}(z|y)), p_{\theta}(z))
\end{equation}

@@

\begin{equation}
    \textup{elbo}_{\lambda_{x}, \lambda_{y}, \beta}(x, \theta_{x}, \theta_{y}, \phi) = E_{q_{\phi}(z|x, \phi)}[\lambda_{x} \textup{ log}(p_{\theta_{x}}(x|z)) + \lambda_{y} \textup{ log}(p_{\theta_{y}}(y|z))] - \beta \textup{ KL}(q_{\phi}(z|x, y), p_{\theta}(z))
\end{equation}

@@

\begin{equation}
    J(x, y, \theta_{x}, \theta_{y}, \phi) = \textup{elbo}_{\lambda_{x}^{1}, \lambda_{y}^{1}, \beta^{1}}(x, y, \theta_{x}, \theta_{y}, \phi) + \textup{elbo}_{\lambda_{x}^{2}, \lambda_{y}^{2}, \beta^{2}}(x, \theta_{x}, \phi) + \textup{elbo}_{\lambda_{x}^{3}, \lambda_{y}^{3}, \beta^{3}}(y, \theta_{y}, \phi)
\end{equation}

@@

@@

@@

\paragraph{Evaluation and Weak Supervision} There is no predictive task for COCO-Captions. Instead we measure the marginal loglikelihood for weak supervision tests.

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage[final]{nips_2017}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{breqn}

@@

\begin{multline}
    J(x, y, \theta, \phi) = \mu (E_{q_{\phi_{x}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{x}}(z|x)), p_{\theta}(z)) \\ + (1 - \mu) (E_{q_{\phi_{y}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{y}}(z|y)), p_{\theta}(z))
\end{multline}

@@

\begin{multline}
    \textup{elbo}_{\lambda_{x}, \lambda_{y}, \beta}(x, \theta_{x}, \theta_{y}, \phi) = E_{q_{\phi}(z|x, \phi)}[\lambda_{x} \textup{ log}(p_{\theta_{x}}(x|z)) + \lambda_{y} \textup{ log}(p_{\theta_{y}}(y|z))] \\ - \beta \textup{ KL}(q_{\phi}(z|x, y), p_{\theta}(z))
\end{multline}

@@

\begin{multline}
    J(x, y, \theta_{x}, \theta_{y}, \phi) = \textup{elbo}_{\lambda_{x}^{1}, \lambda_{y}^{1}, \beta^{1}}(x, y, \theta_{x}, \theta_{y}, \phi) \\ + \textup{elbo}_{\lambda_{x}^{2}, \lambda_{y}^{2}, \beta^{2}}(x, \theta_{x}, \phi) + \textup{elbo}_{\lambda_{x}^{3}, \lambda_{y}^{3}, \beta^{3}}(y, \theta_{y}, \phi)
\end{multline}

@@

\begin{table}[!h]
    \centering
    \begin{tabular}{ l | c | c }
        Model (MNIST) & $\leq \textup{log }p(x)$ & $\leq \textup{log }p(y)$ \\
        \hline
        VAE & & - \\
        BiVCCA & &  \\
        JMVAE & & \\
        MMVAE (Joint) & & \\
        MMVAE (Image-only) & & \\
        MMVAE (Text-only) & & \\
        MMVAE (Anneal KL) & & \\
        MMVAE (With Universal Expert) & & \\
        MMVAE ($\lambda_{y}^{1} = \lambda_{x}^{1} = 0$) & & \\
        MMVAE ($\lambda_{y}^{2} = \lambda_{x}^{3} = 0$) & & \\
        MMVAE ($\lambda_{x}^{2} = \lambda_{y}^{3} = 0$) & & \\
        MMVAE ($\lambda_{x}^{2} = \lambda_{y}^{2} = \lambda_{x}^{3} = \lambda_{y}^{3} = 0$) & & \\
        \newline
    \end{tabular}
    \caption{TODO}
    \label{table:mnist:marginal}
\end{table}

\begin{figure}[!h]
\centering
    \begin{subfigure}[b]{.24\linewidth}
        \centering
        \includegraphics[width=\linewidth]{mnist_mmvae_image_sample.png}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{.24\linewidth}
        \centering
        \includegraphics[width=\linewidth]{mnist_mmvae_condition_on_image_text_9_image_sample.png}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{.24\linewidth}
        \centering
        \includegraphics[width=\linewidth]{mnist_mmvae_condition_on_image_9_image_sample.png}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{.24\linewidth}
        \centering
        \includegraphics[width=\linewidth]{mnist_mmvae_condition_on_text_9_image_sample.png}
        \caption{}
    \end{subfigure}
    \caption{\emph{MNIST Dataset}: MMVAE is bidirectional, allowing us to try samples unconditionally, or conditioned on modalit(ies) taking a particular value. We note that samples conditioned on both image and text are the most sharp and have the lowest variance. As expected, reconstructed images from text alone are the blurriest. (a) Image samples from $z \sim N(0, 1)$. (b) Image samples from $p(z|x=9,y=9)$. (c) Image samples from $p(z|x=9)$. (d) Image samples from $p(z|y=9)$.}
    \label{fig:mnist:samples}
\end{figure}

\begin{table}[!h]
    \centering
    \begin{tabular}{ l | c | c | c | c | c | c | c | c | c | c | c}
        (a) & 5 & 7 & 1 & 3 & 1 & 6 & 3 & 7 & 5 & 6 & 0 \\
        \hline
        (b) & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 \\
        \hline
        (c) & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 \\
        \hline
        (d) & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 \\
        \newline
    \end{tabular}
    \caption{\emph{MNIST Dataset}: Generated text samples from MMVAE trained on MNIST. (a) Text samples from $z \sim N(0, 1)$. (b) Text samples from $p(z|x=9,y=9)$. (c) Text samples from $p(z|x=9)$. (d) Text samples from $p(z|y=9)$. From (b)-(d), we note that conditioning on an image produces the expected text samples.}
    \label{table:mnist:samples}
\end{table}

\begin{figure}[!h]
\centering
    \begin{subfigure}[b]{.32\linewidth}
        \centering
        \includegraphics[width=\linewidth]{mnist_mmvae_pca_manifold.png}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[b]{.32\linewidth}
        \centering
        \includegraphics[width=\linewidth]{mnist_mmvae_joint_weak_supervision.png}
        \caption{}
    \end{subfigure}
    \caption{\emph{MNIST Dataset}: (a) Manifold split by MNIST class; this was generated using the top two principal components from 10k latent samples of dimension 20. We can see that each of the classes form a distinct cluster, suggesting that the latent space is well-defined. (b) The effects of weak supervision on the MNIST prediction task where paired data is only shown a fraction of the time; with 25\% supervision, the model reaches 80\% accuracy.}
    \label{fig:mnist:weaksupervision}
\end{figure}

@@

\begin{table}[!h]
    \centering
    \begin{tabular}{ l | c | c }
        Model (MultiMNIST) & $\leq \textup{log }p(x)$ & $\leq \textup{log }p(y)$ \\
        \hline
        VAE & & - \\
        BiVCCA & &  \\
        JMVAE & & \\
        MMVAE (Joint) & & \\
        MMVAE (Image-only) & & \\
        MMVAE (Text-only) & & \\
        MMVAE (Anneal KL) & & \\
        MMVAE (With Universal Expert) & & \\
        MMVAE ($\lambda_{y}^{1} = \lambda_{x}^{1} = 0$) & & \\
        MMVAE ($\lambda_{y}^{2} = \lambda_{x}^{3} = 0$) & & \\
        MMVAE ($\lambda_{x}^{2} = \lambda_{y}^{3} = 0$) & & \\
        MMVAE ($\lambda_{x}^{2} = \lambda_{y}^{2} = \lambda_{x}^{3} = \lambda_{y}^{3} = 0$) & & \\
        \newline
    \end{tabular}
    \caption{TODO}
    \label{table:multimnist:marginal}
\end{table}

@@

\paragraph{Evaluation and Weak Supervision} There is no predictive task for COCO-Captions. Instead we measure the marginal log likelihood for weak supervision tests.

\begin{table}[!h]
    \centering
    \begin{tabular}{ l | c | c }
        Model (COCO-Caption) & $\leq \textup{log }p(x)$ & $\leq \textup{log }p(y)$ \\
        \hline
        VAE & & - \\
        BiVCCA & &  \\
        JMVAE & & \\
        MMVAE (Joint) & & \\
        MMVAE (Image-only) & & \\
        MMVAE (Text-only) & & \\
        \newline
    \end{tabular}
    \caption{TODO}
    \label{table:mnist:marginal}
\end{table}

\begin{figure}[!h]
    \centering
    \includegraphics[width=.5\linewidth]{coco_mmvae_image_sample.png}
    \caption{This is a work in progress. These samples were generated with a DCGAN encoder and decoder, hence they are very blurry. However, we can make out coherent structure if we squint. I'm hoping that with a PixelCNN decoder (PixelVAE), we will get sharper samples.}
    \label{fig:WIP:coco:samples}
\end{figure}
